cardinalR: Generating interesting high-dimensional data structures

### 2025-05-04 paper-cardinalR.Rmd
Added curvy circle

>>>>

The `gen_curvycycle(n, p)` function generates a $p$-dimensional dataset of $n$ observations lying on a curved closed loop with controlled high-dimensional sinusoidal deviations. The structure forms a one-dimensional nonlinear cycle embedded in higher-dimensional space.

A latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$. The first three dimensions define a non-circular closed curve, referred to as a "curvy cycle." In this configuration, $X_1 = \cos(\theta)$ represents horizontal oscillation, while $X_2 = \sqrt{3}/3 + \sin(\theta)$ introduces a vertical offset to avoid centering the curve at the origin. Additionally, $X_3 = \frac{1}{3} \cdot \cos(3\theta)$ introduces a third harmonic perturbation that intricately folds the curve three times along its path, creating a unique and complex shape that oscillates in both dimensions while incorporating the effects of the harmonic perturbation.

Together, these define a periodic, non-trivial, closed curve in $3\text{-}D$ with internal folds that produce a more complex geometry than a standard circle or ellipse.

For dimensions $X_4$ through $X_p$, additional structured variability is introduced through decreasing amplitude scaling and phase-shifted sine waves. The scaling factor is defined as $\text{scale}_j = \sqrt{(0.5)^{j-3}}$ for $j$ ranging from 4 to $p$, which means that the amplitude decreases as the dimension increases. Each dimension $X_j$ is then calculated using the formula $X_j = \text{scale}_j \cdot \sin(\theta + \phi_j)$, where the phase shift $\phi_j$ is given by $\phi_j = (j - 2) \cdot \frac{\pi}{2p}$.

This introduces decorrelated, low-variance sinusoidal variations across dimensions, retaining the cyclical nature while embedding the data in higher-dimensional space with a coherent but subtle structure.
